Demo arXiv template

### 2022-11-28 paper/paper.qmd
expand dimension reduction section

Sometimes, information from neighbouring spatial locations can also be borrowed when constructing indexes. 
This involves interpolating data from observed locations to the unobserved and aggregating data into different levels.
In the pipeline, spatial aggregation can be written as $$g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),$$ where $\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}$ is the spatial operation parameters and $d_{\theta}$ is the number of parameter of $\theta$. [GIVEN AN EXAMPLE]

collect ground and satellite image -> merge
<!-- $x(s_i^\prime;t_j) = \sum_{i \in \Omega}x(s_i; t_j)$ -->

@@

**Dimension reduction** [TODO]

While multivariate information can be collected on a concept, they need to be summarised into a single number for the final index. Dimension reduction step $$x_{p^*}(\mathbf{s}; \mathbf{t}) \rightarrow x_p(\mathbf{s}; \mathbf{t}),$$ where $p^* = 1, 2, \cdots, P^*$ and $p = 1, 2, \cdots, P$, reduces the variable dimension from $P$ to $P^*$. The most commonly used dimension reduction technique Principal Component Analysis (PCA), also called Empirical Orthogonal Function (EOF) in earth science, can be written as $x_{p^*}(\mathbf{s}; \mathbf{t}) = \sum_{p = 1}^{P}\lambda_{p}x_p(\mathbf{s};\mathbf{t})$ where $\lambda_p$ is the loading of the PC1, derived from maximising the data variance given the constraint $\sum_{p=1}^P\lambda_p^2 = 1$.

weighting

@@

*modell fit? *

Distribution fit can be seen as model fitting in its simplest term. It can be represented by

<!-- With a probability model imposed,  -->

@@

This step maps the univariate series into a different scale, typically for the ease of comparison across regions. For example, a normal scale, [0, 1], or [0, 100] may be favored for reporting certain indexes. In drought indexes, i.e. SPI or SPEI, the quantiles from the fitted distribution is converted into the normal scale via the normal reverse CDF function: $\Phi^{-1}(.)$.

Normalising is usually used in the end of the pipeline and its main difference from the scaling step is that here the change of scale also changes the distribution of the variable. 
While being commonly used, this step can get criticism from analysts for forcing the data into the decided scale, which can be either unnecessary or inaccurately exaggerate or downplay the outliers. Also, the use of normal scale needs to be interpreted with caution. @fig-normalising illustrates the normal density not being directly proportional to its probability of occurrence.
This is concerning, especially at the extreme values, since a small difference in the tail density can have magnitudes of difference in its probability of occurrence.

>>>>

Sometimes

Information from neighbouring spatial locations can also be borrowed when constructing indexes. 
This involves interpolating data from observed locations to the unobserved and aggregating data into different levels.
In the pipeline, spatial aggregation can be written as $$g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),$$ where $\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}$ is the spatial operation parameters and $d_{\theta}$ is the number of parameter of $\theta$. [GIVEN AN EXAMPLE]

While many indexes are calculated independently on spatial locations

satellite images with different resolutions
matching ground and satellite images data

<!-- $x(s_i^\prime;t_j) = \sum_{i \in \Omega}x(s_i; t_j)$ -->

@@

**Dimension reduction**

When constructing indexes from multivariate information, dimension reduction methods combine those information into a univariate series. In the pipeline, dimension reduction includes methods that take multivariate inputs and output the data in a lower dimension (often univariate): $$x_{p^*}(\mathbf{s}; \mathbf{t}) \rightarrow x_p(\mathbf{s}; \mathbf{t}),$$ where $p^* = 1, 2, \cdots, P^*$ and $p = 1, 2, \cdots, P$ reduce the variable dimension from $P$ to $P^*$. 
The most commonly used dimension reduction technique is Principal Component Analysis (PCA), also called Empirical Orthogonal Function (EOF) in earth science. It can be written as $x_{p^*}(\mathbf{s}; \mathbf{t}) = \sum_{p = 1}^{P}\lambda_{p}x_p(\mathbf{s};\mathbf{t})$ where $\lambda_p$ is the loading of the PC1, derived from maximising the data variance given the constraint $\sum_{p=1}^P\lambda_p^2 = 1$.

The PCA is a special case of weighting, where individual variables are summed up in a linear combination with potential restrictions imposed on the weight coefficients. These coefficients may be simple mathematical constructions that give equal weight to each variable or can be derived with specialised knowledge by experts in the field. 
While suggested weights can indicate norms adored by practitioners, analysts should be given the flexibility to experiment with different combinations when constructing indexes. 
This could help understand index behavior from its sensitivity to the variables and suggest alternative weights that better suit specific tasks.

@@

*model fit? *

Distribution fit can be seen as model fitting in its simplest term. It can be represented by 
<!-- With a probability model imposed,  -->

@@

This step maps the univariate series into a different scale, typically for the ease of comparison across regions. For example, a normal scale, [0, 1], or [0, 100] may be favored for reporting certain indexes. In drought indexes, i.e. SPI or SPEI, the quantiles from the fitted distribution is converted into the normal scale via the normal reverse CDF function: $\Phi^{-1}(.)$. Normalising is usually used in the end of the pipeline and its main difference from the scaling step is that here the change of scale also changes the distribution of the variable. 
While being commonly used, this step can get criticism from analysts for forcing the data into the decided scale, which can be either unnecessary or inaccurately exaggerate or downplay the outliers. Also, the use of normal scale needs to be interpreted with caution. @fig-normalising illustrates the normal density not being directly proportional to its probability of occurrence.
This is concerning, especially at the extreme values, since a small difference in the tail density can have magnitudes of difference in its probability of occurrence.

### 2022-11-28 paper/paper.tex
expand dimension reduction section

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, interior hidden]}{\end{tcolorbox}}\fi

@@

Sometimes, information from neighbouring spatial locations can also be
borrowed when constructing indexes. This involves interpolating data
from observed locations to the unobserved and aggregating data into
different levels. In the pipeline, spatial aggregation can be written as
\[g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),\] where
\(\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}\) is the spatial
operation parameters and \(d_{\theta}\) is the number of parameter of
\(\theta\). {[}GIVEN AN EXAMPLE{]}

collect ground and satellite image -\textgreater{} merge

@@

\textbf{Dimension reduction} {[}TODO{]}

While multivariate information can be collected on a concept, they need
to be summarised into a single number for the final index. Dimension
reduction step
\[x_{p^*}(\mathbf{s}; \mathbf{t}) \rightarrow x_p(\mathbf{s}; \mathbf{t}),\]
where \(p^* = 1, 2, \cdots, P^*\) and \(p = 1, 2, \cdots, P\), reduces
the variable dimension from \(P\) to \(P^*\). The most commonly used
dimension reduction technique Principal Component Analysis (PCA), also
called Empirical Orthogonal Function (EOF) in earth science, can be
written as
\(x_{p^*}(\mathbf{s}; \mathbf{t}) = \sum_{p = 1}^{P}\lambda_{p}x_p(\mathbf{s};\mathbf{t})\)
where \(\lambda_p\) is the loading of the PC1, derived from maximising
the data variance given the constraint \(\sum_{p=1}^P\lambda_p^2 = 1\).

weighting

@@

\emph{modell fit? }

@@

This step maps the univariate series into a different scale, typically
for the ease of comparison across regions. For example, a normal scale,
{[}0, 1{]}, or {[}0, 100{]} may be favored for reporting certain
indexes. In drought indexes, i.e.~SPI or SPEI, the quantiles from the
fitted distribution is converted into the normal scale via the normal
reverse CDF function: \(\Phi^{-1}(.)\).

Normalising is usually used in the end of the pipeline and its main
difference from the scaling step is that here the change of scale also
changes the distribution of the variable. While being commonly used,
this step can get criticism from analysts for forcing the data into the
decided scale, which can be either unnecessary or inaccurately
exaggerate or downplay the outliers. Also, the use of normal scale needs
to be interpreted with caution. Figure~\ref{fig-normalising} illustrates
the normal density not being directly proportional to its probability of
occurrence. This is concerning, especially at the extreme values, since
a small difference in the tail density can have magnitudes of difference
in its probability of occurrence.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

@@

Sometimes

Information from neighbouring spatial locations can also be borrowed
when constructing indexes. This involves interpolating data from
observed locations to the unobserved and aggregating data into different
levels. In the pipeline, spatial aggregation can be written as
\[g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),\] where
\(\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}\) is the spatial
operation parameters and \(d_{\theta}\) is the number of parameter of
\(\theta\). {[}GIVEN AN EXAMPLE{]}

While many indexes are calculated independently on spatial locations

satellite images with different resolutions matching ground and
satellite images data

@@

\textbf{Dimension reduction}

When constructing indexes from multivariate information, dimension
reduction methods combine those information into a univariate series. In
the pipeline, dimension reduction includes methods that take
multivariate inputs and output the data in a lower dimension (often
univariate):
\[x_{p^*}(\mathbf{s}; \mathbf{t}) \rightarrow x_p(\mathbf{s}; \mathbf{t}),\]
where \(p^* = 1, 2, \cdots, P^*\) and \(p = 1, 2, \cdots, P\) reduce the
variable dimension from \(P\) to \(P^*\). The most commonly used
dimension reduction technique is Principal Component Analysis (PCA),
also called Empirical Orthogonal Function (EOF) in earth science. It can
be written as
\(x_{p^*}(\mathbf{s}; \mathbf{t}) = \sum_{p = 1}^{P}\lambda_{p}x_p(\mathbf{s};\mathbf{t})\)
where \(\lambda_p\) is the loading of the PC1, derived from maximising
the data variance given the constraint \(\sum_{p=1}^P\lambda_p^2 = 1\).

The PCA is a special case of weighting, where individual variables are
summed up in a linear combination with potential restrictions imposed on
the weight coefficients. These coefficients may be simple mathematical
constructions that give equal weight to each variable or can be derived
with specialised knowledge by experts in the field. While suggested
weights can indicate norms adored by practitioners, analysts should be
given the flexibility to experiment with different combinations when
constructing indexes. This could help understand index behavior from its
sensitivity to the variables and suggest alternative weights that better
suit specific tasks.

@@

\emph{model fit? }

@@

This step maps the univariate series into a different scale, typically
for the ease of comparison across regions. For example, a normal scale,
{[}0, 1{]}, or {[}0, 100{]} may be favored for reporting certain
indexes. In drought indexes, i.e.~SPI or SPEI, the quantiles from the
fitted distribution is converted into the normal scale via the normal
reverse CDF function: \(\Phi^{-1}(.)\). Normalising is usually used in
the end of the pipeline and its main difference from the scaling step is
that here the change of scale also changes the distribution of the
variable. While being commonly used, this step can get criticism from
analysts for forcing the data into the decided scale, which can be
either unnecessary or inaccurately exaggerate or downplay the outliers.
Also, the use of normal scale needs to be interpreted with caution.
Figure~\ref{fig-normalising} illustrates the normal density not being
directly proportional to its probability of occurrence. This is
concerning, especially at the extreme values, since a small difference
in the tail density can have magnitudes of difference in its probability
of occurrence.
